---
title: "Data Consolidation and Cleaning"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

{r setup, include=FALSE}

## Data Cleaning and Consolidation

## There were 2 rounds of data cleaning and consolidation involved. First round happened when each members have prepared and cleaned each individual ##datasets for each domain. After the cleaned dataset was passed on to other group members, some follow-up for data cleaning and consolidation was ##done. 


# 1. Read Libraries
```{r}
library("readxl")
library(tidyverse)
library(dplyr)
```


# 2. Load all datasets 

Education 
```{r}
education_latest_data <- as.data.frame(read_excel("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/Education_Datasets_Summary.xlsx", sheet = "Latest_Available_Data"))

education_5YearsMean_data <- as.data.frame(read_excel("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/Education_Datasets_Summary.xlsx", sheet = "Past_5_Years_Mean"))

```



Health
```{r}
health_latest_data <- as.data.frame(read_excel("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/HealthData_MergedNew.xlsx", sheet = "LatestAvailableData"))

health_5YearsMean_data <- as.data.frame(read_excel("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/HealthData_MergedNew.xlsx", sheet = "Past5YrsMean"))
```




Health Facilities - Data Type Conversion was made to some variables here 
```{r}

# Health Facilities

health_fac_latest_data <- as.data.frame(read_excel("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/LW_Datasets_Merged.xlsx", sheet = "Latest"))

health_fac_5YearsMean_data <- as.data.frame(read_excel("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/LW_Datasets_Merged.xlsx", sheet = "Mean"))



# Transform the chr datatype into numeric

# latest data

health_fac_latest_data_convert <- health_fac_latest_data %>% 
  select(-c('Country_Name', 'Country_Code')) %>% 
  mutate_if(is.character, as.numeric)

glimpse(health_fac_latest_data_convert)

health_fac_latest_data <- cbind(health_fac_latest_data[,c('Country_Name', 'Country_Code')], health_fac_latest_data_convert)


glimpse(health_fac_latest_data)

# 5 years mean

health_fac_5YearsMean_data_convert <- health_fac_5YearsMean_data %>% 
  select(-c('Country_Name', 'Country_Code')) %>% 
  mutate_if(is.character, as.numeric)

glimpse(health_fac_5YearsMean_data_convert)

health_fac_5YearsMean_data <- cbind(health_fac_5YearsMean_data[,c('Country_Name', 'Country_Code')], health_fac_5YearsMean_data_convert)


glimpse(health_fac_5YearsMean_data)

```



General Data
```{r}
general_data_latest <- as.data.frame(read_excel("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/Ricky-Data-Summary.xlsx", sheet = "Latest"))
general_data_5YearsMean_data <- as.data.frame(read_excel("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/Ricky-Data-Summary.xlsx", sheet = "5_Year_Mean"))
```



Covid Data
```{r}
covid_data <- read.csv("https://raw.githubusercontent.com/datasets/covid-19/master/data/countries-aggregated.csv")

latest_covid <- filter(covid_data,Date == '2020-12-19') #Latest Available Data

# view first 5 rows of covid data
head(latest_covid)
```




# 3.Joining / Merging tables

5 Year Mean Data
```{r}

### Combining Latest Covid Date with 5-Year Mean Data ###


# 1. Left Join Education to the covid data

mean_data_1 <- left_join(latest_covid, education_5YearsMean_data, by = c("Country" = "Country Name")) 

# 2. Left join Health to 1.

mean_data_2 <- left_join(mean_data_1, health_5YearsMean_data, by = c("Country" = "Country Name")) 


# 3. Left join General Data to 2.

mean_data_3 <- left_join(mean_data_2, general_data_5YearsMean_data, by = c("Country" = "Country Name")) 


# 4. Left join General Data to 3.

mean_data_4 <- left_join(mean_data_3, health_fac_5YearsMean_data, by = c("Country Code.x" = "Country_Code")) 


# Drop repeated & Unnecessary Columns

glimpse(mean_data_4)


mean_data_dropped_redundant_cols <- mean_data_4 %>%  select(-c('Country Code.x', 'Country Code.y', 'Country_Name'))


glimpse(mean_data_dropped_redundant_cols)



# Check missing values of each variables (columns)

mean_data_view_missing_data <- colSums(is.na(mean_data_dropped_redundant_cols) == TRUE)

glimpse(mean_data_view_missing_data)

```



5 Year Latest Data (2019)
```{r}

### Combining Latest Covid Date with the latest Data ###


# 1. Left Join Education to the covid data

latest_data_1 <- left_join(latest_covid, education_latest_data, by = c("Country" = "Country Name")) 


# 2. Left join Health to 1.

latest_data_2 <- left_join(latest_data_1, health_latest_data, by = c("Country" = "Country Name")) 



# 3. Left join General Data to 2.

latest_data_3 <- left_join(latest_data_2, general_data_latest, by = c("Country" = "Country Name")) 



# 4. Left join General Data to 3.

latest_data_4 <- left_join(latest_data_3, health_fac_latest_data, by = c("Country Code.x" = "Country_Code")) 



# Drop repeated & Unnecessary Columns

glimpse(latest_data_4)


latest_data_dropped_redundant_cols <- latest_data_4 %>%  select(-c('Country Code.x', 'Country Code.y', 'Country Code', 'Country_Name'))


glimpse(latest_data_dropped_redundant_cols)



# Check missing values of each variables (columns)

latest_data_view_missing_data <- colSums(is.na(latest_data_dropped_redundant_cols) == TRUE)

glimpse(latest_data_view_missing_data)
```




# Findings: There are more data for LATEST DATASET, hence the analysis will be done using latest dataset
```{r}


glimpse(latest_data_dropped_redundant_cols)

analysis_dataset <- latest_data_dropped_redundant_cols
```


# 4. Save Dataset
```{r}

### Save ###

library("writexl")
write_xlsx(analysis_dataset,"consolidate_dataset.xlsx")

```




# Follow Up - 2nd round of cleaning




# Merge latest dataset
```{r}

latest_data_combined <- merge(x = education_latest_data, y = health_latest_data, by = "Country Code")

latest_data_combined2 <- merge(x = latest_data_combined, y = general_data_latest, by = "Country Code")

latest_data_combined3 <- merge(x = latest_data_combined2, y = health_fac_latest_data, by.x = "Country Code", by.y = "Country_Code")

glimpse(latest_data_combined3)
```


# Drop redundant data
```{r}


latest_data_combined4 <- latest_data_combined3 %>% select(-c("Country Name.x", "Country Name.y", "Country_Name", "Country Name"))

names(latest_data_combined4)[names(latest_data_combined4) == 'Country Code'] <- "Country_Code"


glimpse(latest_data_combined4)
```



# Read continents data 
```{r}

continent <- read.csv("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets//Cleaning/continents2.csv")

head(continent)


continent2 <- continent %>% select(c('ï..name', alpha_3, region, sub_region, intermediate_region))
 
 names(continent2)[names(continent2) == "ï..name"] <- "Country_Name"
 names(continent2)[names(continent2) == "alpha_3"] <- "Country_Code"
 names(continent2)[names(continent2) == "region"] <- "Region"
 names(continent2)[names(continent2) == "sub_region"] <- "Sub_Region"
 names(continent2)[names(continent2) == "intermediate_region"] <- "Intermediate_Region"



glimpse(latest_covid)

glimpse(continent2)

```




# Merge Continent and Covid Data
```{r}

glimpse(latest_covid)

glimpse(continent2)

DFCovid2 <- merge(x = continent2, y = latest_covid, by.x = "Country_Name", by.y = "Country",all.y = TRUE) # right join


#remove date column
DFCovid3 <- DFCovid2 %>% select(-c(Date))


glimpse(DFCovid3)

```


# Join df with aggregated data and Manual Preprocessing for US country name
```{r}

# import samuel data
agg_data <- read.csv("C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/Cleaning/consolidate_dataset_edited_FilteredCSV3.csv")

agg_data2 <- agg_data %>% select(c(code,ConfirmedPerPopulation...,RecoveryRate...,DeathRate...))

glimpse(agg_data2)

names(agg_data2)[names(agg_data2) == "code"] <- "Country_Code"
names(agg_data2)[names(agg_data2) == "ConfirmedPerPopulation..."] <- "ConfirmedPerPpopulation"
names(agg_data2)[names(agg_data2) == "RecoveryRate..."] <- "RecoveryRate"
names(agg_data2)[names(agg_data2) == "DeathRate..."] <- "DeathRate"


glimpse(agg_data2)


Covid_Rate_Data <- merge(x = DFCovid3, y = agg_data2, by.x = "Country_Code", by.y = "Country_Code",all.x = TRUE)

# Change the Country_Name , US to United States

Covid_Rate_Data[193, 2] <- 'United States'
Covid_Rate_Data[193, 1] <- 'USA'

head(Covid_Rate_Data)
```



# Combine the latest data with covid data and write data into csv
```{r}

glimpse(Covid_Rate_Data)

latest_data <- merge(x = Covid_Rate_Data, y = latest_data_combined4, by.x = "Country_Code", by.y = "Country_Code",all.x = TRUE)



latest_data_ver2 <- unique(latest_data)


# check how many repeated countries

unique_country <- unique(latest_data$Country_Code)
dim(unique_country)

unique_country2 <- unique(latest_data$Country_Name)
dim(unique_country2)


write.csv(latest_data_ver2,"C:/Users/uwxl003/Downloads/PoDS - Consolidate Datasets/PoDS - Consolidate Datasets/Cleaning/latest_cleansed_data.csv", row.names = FALSE)
```